---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

[![Build Status](
  https://travis-ci.org/russHyde/dupree.svg?branch=master
  )
](
  https://travis-ci.org/russHyde/dupree
)

# dupree

The goal of `dupree` is to identify chunks / blocks of highly duplicated code within a set of R scripts.

A very lightweight approach is used:

- The user provides a set of `*.R` and/or `*.Rmd` files;

- All R-code in the user-provided files is read and code-blocks are identified;

- The non-trivial symbols from each code-block are retained (for instance, really common symbols like `<-`, `,`, `+`, `(` are dropped);

- Similarity between different blocks is calculated using stringdist::seq_sim by longest-common-subsequence (symbol-identity is at whole-word level - so "my_data", "my_Data", "my.data" and "myData" are not considered to be identical in the calculation - and all non-trivial symbols have equal weight in the similarity calculation);

- Code-blocks pairs (both between and within the files) are returned in order of highest similarity

To prevent the results being dominated by high-identity blocks containing very few symbols (eg, `library(dplyr)`) the user can specify a `min_block_size`. Any code-block containing at least this many non-trivial symbols will be kept.

## Installation

You can install `dupree` from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("russHyde/dupree")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
## basic example code
library(dupree)
files <- dir(pattern = "*.R(md)*$", recursive = TRUE)
dupree(files, min_block_size = 20)
```
