---
title: "Clean(ish) Code: `lintr` and `dupree`"
subtitle: "Presented to EdinbR R-Users Group, 2019-07-17"
author: "Russ Hyde, University of Glasgow"
output:
  ioslides_presentation:
    df_print: paged
---

<!--
  Styling to allow footnotes to appear on the same slide as their definition
  - Borrowed from https://stackoverflow.com/questions/42690955
-->

<style>
div.footnotes {
  position: absolute;
  bottom: 0;
  margin-bottom: 10px;
  width: 80%;
  font-size: 0.6em;
}
</style>

<script
  src="https://ajax.googleapis.com/ajax/libs/jquery/3.1.1/jquery.min.js"
  ></script>
<script>
  $(document).ready(function() {
    $('slide:not(.backdrop):not(.title-slide)').append('<div
class=\"footnotes\">');

    $('footnote').each(function(index) {
      var text  = $(this).html();
      var fnNum = (index+1).toString().sup();
      $(this).html(text + fnNum);

      var footnote   = fnNum + ': ' + $(this).attr('content') + '<br/>';
      var oldContent =
$(this).parents('slide').children('div.footnotes').html();
      var newContent = oldContent + footnote;
      $(this).parents('slide').children('div.footnotes').html(newContent);
    });
  });
</script>

<!-- Packages -->

```{r, echo = FALSE}
# suppressPackageStartupMessages({
#   library(lintr)
# })
```

<!-- Presentation code -->

## Background and Links:

### Today's packages

- [https://github.com/russHyde/dupree](https://github.com/russHyde/dupree)
- [https://github.com/jimhester/lintr](https://github.com/jimhester/lintr)

### Me

- [https://twitter.com/haematobot](https://twitter.com/haematobot)
- [https://russ-hyde.rbind.io/](https://russ-hyde.rbind.io/)
- [Paul O'Gorman Leukaemia Research Centre](
    https://www.gla.ac.uk/connect/supportus/medicalfund/paulogormanleukaemiaresearchcentre/
  )

## Preamble

```{r}
# Dependencies:
# - lintr, dplyr, purrr, tibble, magrittr, methods, stringdist
if (!"dupree" %in% installed.packages()) {
  require("devtools")
  devtools::install_github(
    repo = "russHyde/dupree", dependencies = FALSE
  )
}

suppressPackageStartupMessages({
  library(lintr)
  library(dupree)
})
```
