---
title: "Clean(ish) Code: `dupree`"
subtitle: "Presented to EdinbR R-Users Group, 2019-07-17"
author: "Russ Hyde, University of Glasgow"
output:
  ioslides_presentation:
    df_print: paged
---

<!--
  Styling to allow footnotes to appear on the same slide as their definition
  - Borrowed from https://stackoverflow.com/questions/42690955

  To include a footnote use the syntax
  Blah Blah Blah <footnote content="xyz"> Blah </footnote>
-->

<style>
div.footnotes {
  position: absolute;
  bottom: 0;
  margin-bottom: 10px;
  width: 80%;
  font-size: 0.6em;
}
</style>

<script src="https://ajax.googleapis.com/ajax/libs/jquery/3.1.1/jquery.min.js"></script>

<script>
  $(document).ready(function() {
    $('slide:not(.backdrop):not(.title-slide)').append('<div class=\"footnotes\">');

    $('footnote').each(function(index) {
      var text  = $(this).html();
      var fnNum = (index+1).toString().sup();
      $(this).html(text + fnNum);

      var footnote   = fnNum + ': ' + $(this).attr('content') + '<br/>';
      var oldContent = $(this).parents('slide').children('div.footnotes').html();
      var newContent = oldContent + footnote;
      $(this).parents('slide').children('div.footnotes').html(newContent);
    });
  });
</script>

<!-- Packages -->

```{r, echo = FALSE}
suppressPackageStartupMessages({
  library(dplyr)
  library(magrittr)
  library(tidygraph)
  library(ggraph)
  library(stringdist)
})
```

<!-- Presentation code -->

## Background and Links:

### Today's packages

- [https://github.com/russHyde/dupree](https://github.com/russHyde/dupree)
- [https://github.com/jimhester/lintr](https://github.com/jimhester/lintr)
- [`https://github.com/ropensci/git2r`](https://github.com/ropensci/git2r)

### Me

- [https://twitter.com/haematobot](https://twitter.com/haematobot)
- [https://russ-hyde.rbind.io/](https://russ-hyde.rbind.io/)
- [Paul O'Gorman Leukaemia Research Centre](
    https://www.gla.ac.uk/connect/supportus/medicalfund/paulogormanleukaemiaresearchcentre/
  )

## Preamble

```{r}
# Dependencies:
# - lintr, dplyr, purrr, tibble, magrittr, methods, stringdist
if (!"dupree" %in% installed.packages()) {
  require("devtools")
  devtools::install_github(
    repo = "russHyde/dupree", dependencies = FALSE
  )
}

suppressPackageStartupMessages({
  library(lintr)
  library(dupree)
  library(git2r)
})
```

## Code Smells & Architectural Ideals

_"The most common design problems result from code that_

- _Is duplicated_

- _Is unclear_

- _Is complicated"_

Quote: Kerievsky 'Refactoring to Patterns'

See also
Fowler 'Refactoring',
Martin 'Clean Code'
and Jenny Bryan's talk 'Code smells and feels'

## Types of duplication

```{r, echo = FALSE}
url <- "https://static.fjcdn.com/large/pictures/9d/b7/9db733_1672275.gif"
```

![](`r url`){ width=80% }

- Trivial stuff (`library(dplyr)`)

- Copy/paste-driven development (similar logic & code)

- Functional duplication (same logic, different code)

- ? False duplication (different logic, similar code)

## How to detect duplication?

- Python
    - `pylint` (looks for identical _lines_ between files)

- Java / C++ / C# etc
    - lots of choice (code structure / identity)

- R: nothing for source code (AFAIK)

    - String / Sequence similarity: `stringdist`
    - Text analysis: `ropensci:textreuse`
    - (But tools like: `goodpractice`, `lintr`, `styler`, `cyclocomp`, `pkgnet`)

## `dupree`

- [https://github.com/russHyde/dupree](https://github.com/russHyde/dupree)

- All community input is welcome

- Most data input is welcome:
    - sets of files (`dupree()`)
    - a directory (`dupree_dir()`)
    - or a package (`dupree_package()`)

## Mechanics

Longest Common Substring
```{r}
# breakf-a---st
# break-dance--
stringdist::stringdist("breakfast", "breakdance", method = "lcs")
```
Code blocks

-> Sentences of function / variable names

-> "Sentences" of integers

-> Compute similarity score based on longest-common-subsequence

## Mechanics (cont.)

Use `seq_sim` to compute LCS-based distance between vectors of integers

```{r}
to_ints <- function(word){
  as.integer(factor(strsplit(word, "")[[1]], levels = letters))
}

to_ints("breakfast")
```

```{r}
stringdist::seq_sim(
  list(to_ints("breakfast")), list(to_ints("breakdance")), method = "lcs"
) # 1 - |LCS| / (|seq1| + |seq2|)
```

## Duplication in a script

```{r}
# min_block_size: used to prevent dupree analysing really small code blocks
dupree("duplication_heavy.R", min_block_size = 3) %>%
  dplyr::select(-file_a, -file_b)
```

## Duplication in a script (cont.) {.smaller}

```{r, code = readLines("duplication_heavy.R"), eval = FALSE}
```

<!--
## Duplication in a script {.smaller}

```{r, echo = FALSE}
temp_lear <- tempfile(fileext = ".R")
```

```{r, echo = FALSE}
lear <- "
# temp file

lear1 <- function() {
  print(paste('there', 'was', 'an', 'old', 'man' 'of', 'Cashmere'))
  print(paste('whose', 'movements', 'were', 'scroobious', 'and', 'queer'))
  print(paste('being', 'slender', 'and', 'tall', 'he', 'looked', 'over', 'a', 'wall')),
  print(paste('and', 'perceived', 'two', 'fat', 'ducks', 'of', 'Cashmere'
  ))
}

lear2 <- function(){
  paste(
    'there', 'was', 'an', 'old', 'man' 'of', 'black', 'heath',
    'whose', 'head', 'was', 'adorned', 'with', 'a', 'wreath',
    'of', 'lobsters', 'and', 'spice', 'pickled', 'onions', 'and' 'mice',
    'that', 'uncommon', 'old', 'man', 'of', 'black', 'heath'
  )
}

citation('Lear: Complete Nonsense')
"

cat(lear, file = temp_lear)
```

```{r, code = readLines(temp_lear), eval = FALSE}
```

## Duplication in a script (cont.)

```{r}
dupree(file = temp_lear, min_block_size = 3)
```

-->

## Duplication in a package

Downloaded the source code for `lintr` from github using
[`ropensci/git2r`](https://github.com/ropensci/git2r).

```{r, message = FALSE, results="hide"}
# temporary dir for storing `lintr`'s source code
lintr_path <- file.path(tempdir(), "lintr")
lintr_repo <- git2r::clone(
  "https://github.com/jimhester/lintr",
  lintr_path
)
```

## Duplication in a package (cont)

Ran dupree on `lintr`

```{r}
dups <- dupree::dupree_package(
  lintr_path, min_block_size = 40
)
```

```{r, fig.height=3, echo = FALSE}
ggplot2::qplot(
  x = seq(nrow(dups)), y = dups[["score"]],
  xlab = "Index",
  ylab = "Similarity score:\nversus closest matching block")
```

## Duplication in a package (cont) {.smaller}

```{r}
dups %>%
  dplyr::filter(score > 0.4 & file_a != file_b) %>%
  dplyr::mutate_at(c("file_a", "file_b"), basename) %>%
  head()
```

[GOTO: `equals_na_lintr.R`](https://github.com/jimhester/lintr/blob/master/R/equals_na_lintr.R)

## Visualisation of duplication results

We make a tidygraph structure from the similarity scores

```{r}
dup_graph <- dups %>%
  # keep code-block pairs with moderate similarity:
  dplyr::filter(score > 0.4) %>%
  dplyr::transmute(
    # indicate code-blocks by filename and start-line
    from = paste(basename(file_a), line_a),
    to = paste(basename(file_b), line_b),
    type = "duplication",
    score = score
  ) %>%
  tidygraph::as_tbl_graph() %>%
  # distinguish the file each code block came from
  mutate(filename = gsub("(.*) \\d+$", "\\1", name))
```

<!--

## graph

```{r}
dup_filtered <- filter(dups, score > 0.4)

sequence_graph <- rbind(
  transmute(dup_filtered, file = basename(file_a), line = line_a),
  transmute(dup_filtered, file = basename(file_b), line = line_b)
  ) %>%
  unique() %>%
  group_by(file) %>%
  # dplyr::group_map requires dplyr >= 0.8
  group_map(function(.x, .y) {
    ordered <- arrange(.x, line)
    tibble::tibble(
      file = .y$file,
      from = c("start", ordered$line),
      to = c(ordered$line, "end")
    ) %>%
      transmute(
        from = paste(file, from),
        to = paste(file, to),
        type = "sequence",
        score = 0
      )
  }) %>%
  bind_rows()
```

## graph

```{r}
seq_graph <- sequence_graph %>%
  as_tbl_graph() %>%
  mutate(
    filename = gsub("(.*) \\d+$", "\\1", name),
    y = as.integer(gsub("(.*) (\\d+)$", "\\2", name)),
    x = as.integer(as.factor(filename))
  )
```

```{r, echo = FALSE}
bind_graphs(
  seq_graph, dup_graph
) %>%
  ggraph(layout = "gem") +
  geom_edge_link(
    aes(colour = type, edge_width = score)
  ) +
  geom_node_point(
    aes(colour = filename), size = 4, show.legend = FALSE
  ) +
  theme_graph()
```

-->

## Visualisation of duplication results (cont)

```{r}
graph_image <- dup_graph %>%
  ggraph(layout = "gem") +
  geom_edge_link(
    aes(colour = type, edge_width = score)
  ) +
  geom_node_point(
    aes(colour = filename), size = 4, show.legend = FALSE
  ) +
  theme_graph()
```

## Visualisation of duplication results (cont)

```{r}
graph_image
```

## Visualisation of duplication results (cont)

```{r}
graph_image +
  geom_node_text(aes(label = name), repel = TRUE)
```

## What was `lintr` by the way?

- style / syntax checker for _R_

- configurable

- can be ran

    - in Rstudio / vim / atom etc

    - or on Travis

- (and dupree uses lintr's file parsers)

## refactoRing

- Improving the structure of code (without modifying its function)

- The rule of 3

- Examples

    - Figures: Global theming / %+%

    - Statements: Replace with function call

    - Common functions: Move to a package

    - RMarkdown: Configurable reports / child-stubs

# Thanks
